Poker Chips, Earthquakes, and Computation

### 2014-09-26 draft0.tex
edited some of the definitions and notation section

\begin{definition} A board is a map from the integer lattice to the non-negative integers.
That is a Board $f$ is an element of $\mathcal{B} := \{ f: \mathbb{Z} \times \mathbb{Z} \rightarrow \mathbb{Z}_{\geq 0} | f \text{ is a function }\}$.
The input values of a board represent the coordinates on the integer lattice that the lower left hand corner of square occupies and the output value refers to the number of chips on that square. 
A square is an element of $\mathbb{Z} \times \mathbb{Z}$. 
\end{definition}

\begin{definition} We define the global earthquake operator as a map from the set of all boards to the set of all boards and write $GE: \mathcal{B} \rightarrow \mathcal{B}$. 
It transforms all of the squares on the board with the following rule.
\begin{equation}
GE(f(x,y)) = f(x,y) - 4[[f(x,y) \geq 4]] + [[f(x-1,y) \geq 4]] + [[f(x+1,y) \geq 4]] + [[f(x,y-1) \geq 4]] + [[f(x,y+1) \geq 4]],
\end{equation}
where $[[A]]$ is defined as 
\begin{equation}
[[A]] = 
\begin{cases} 
1 \text{ if } A \text{ is true} \\ 
0 \text{ if } A  \text{ is false}
 \end{cases}.
\end{equation}
\end{definition}

\begin{definition}
 We define the local earthquake operator as a map from the set of all boards and locations to the set of all boards and write $LE: \mathcal{B} \times (\mathbb{Z} \times \mathbb{Z}) \rightarrow \mathcal{B}$.
If $f(x,y) \leq 4$, then $LE(f, (x,y)) = f$, otherwise,  
\begin{align*}
LE( f(x,y), (x,y)) = f(x,y)-4 \\
LE( f(x+1,y), (x,y)) = f(x+1,y)+1 \\
LE( f(x-1,y), (x,y)) = f(x-1,y)+1 \\
LE( f(x,y+1), (x,y)) = f(x,y+1)+1 \\
LE( f(x,y-1), (x,y)) = f(x,y-1)+1,
\end{align*}
and leaves all other squares unchanged.

We use the notation $LE^n(f, \{(x_i,y_i)\})$ to refer to the application of $LE$ to the board $f$, $n$ times to the squares $(x_1, y_1), \ldots (x_{n}, y_{n})$ respectively. 
\end{definition}

>>>>

First, we formalize the board of poker chips. We can overlay the grid on the $xy$ plane so that each grid square is a $1$ by $1$ square, and each grid intersection point is at a coordinate $(a,b)$ where $a,b \in \mathbb{Z}$. Then, the coordinate of a given grid square is just the coordinate of the grid intersection point of its lower left corner.

\begin{definition} A board is a function $f: \mathbb{Z} \times \mathbb{Z} \to \mathbb{Z}_{\geq 0}$.
The inputs to this function correspond to coordinates of grid squares and the output is the number of chips on that square. Let $\mathcal{B}$ denote the set of all boards.
\end{definition}

Recall from the introduction that local earthquakes (abbreviated as $LE$) are operations that hit one square with $\geq 4$ chips and redistributes 1 chip to each adjacent square. We define this operation in terms of $f$.

\begin{definition}
A local earthquake takes as input a board and the coordinates of a grid square and outputs a board, so we write $LE: \mathcal{B} \times (\mathbb{Z} \times \mathbb{Z}) \rightarrow \mathcal{B}$. If $f(x,y) \geq 4$,
\begin{align*}
LE( f(x,y), (x,y)) = f(x,y)-4 \\
LE( f(x+1,y), (x,y)) = f(x+1,y)+1 \\
LE( f(x-1,y), (x,y)) = f(x-1,y)+1 \\
LE( f(x,y+1), (x,y)) = f(x,y+1)+1 \\
LE( f(x,y-1), (x,y)) = f(x,y-1)+1,
\end{align*}
Otherwise if $f(x,y) \leq 4$, then $LE(f(x,y)) = f$, as the earthquake does not change the board.
\end{definition}
We use the notation $LE^n(f, \{(x_i,y_i)\})$ to denote the application of $LE$ to the board $f$, $n$ times to the squares $(x_1, y_1), \ldots (x_{n}, y_{n})$ respectively.

Recall that a global earthquake (abbreviated as $GE$) is equivalent to a local earthquake hitting every square on the board.

\begin{definition} We define the global earthquake operator as a map from the set of all boards to the set of all boards and write $GE: \mathcal{B} \rightarrow \mathcal{B}$. 
It transforms all of the squares on the board with the following rule.
\begin{align*}
GE(f(x,y)) = &f(x,y) - 4[[f(x,y) \geq 4]] + [[f(x-1,y) \geq 4]] + [[f(x+1,y) \geq 4]] + \\
&[[f(x,y-1) \geq 4]] + [[f(x,y+1) \geq 4]],
\end{align*}
where $[[A]]$ is defined as 
\begin{equation}
[[A]] = 
\begin{cases} 
1 \text{ if } A \text{ is true} \\ 
0 \text{ if } A  \text{ is false}
 \end{cases}.
\end{equation}
\end{definition}
